Undirected Graph Exploration with $\Theta(\log\log n)$ Pebbles

### 2016-05-12 presentation/main.tex
slides!

\usepackage[utf8]{inputenc}
\usepackage{epigraph}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{mdframed}
\usepackage{graphicx}
\usepackage{appendixnumberbeamer}

@@

\begin{frame}
  \frametitle{Agents}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Agent carries set of distinguishable pebbles
      \item Formalisation as pebble machines
        \begin{itemize}
          \item $(s,p,m)$-pebble machine
          \item $(Q,F,P,m,\delta,\din,\dout,q_{0})$
          \item<2-> $\din$: moving onto vertex
          \item<8-> $\dout$: leaving vertex
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \begin{itemize}
      \item<3-> As an agent moves onto a vertex it observes:
        \begin{itemize}
          \item<4-> Port it entered through
          \item<5-> Pebbles it carries
          \item<6-> Pebbles on the vertex
          \item<7-> Degree of vertex
        \end{itemize}
      \item<9-> Computation on a vertex determines:
        \begin{itemize}
          \item<10-> Port to leave
          \item<11-> Pebbles to drop
          \item<12-> Pebbles to carry along
        \end{itemize}
    \end{itemize}
  \end{columns}
\end{frame}

@@

\begin{frame}
  \frametitle{$3$-regularity}
  \begin{mdframed}[linecolor=purple]
    Establish $3$-regularity
  \end{mdframed}
  \begin{center}
    \uncover<2->{\resizebox{0.7\textwidth}{!}{\input{tikz/regsim}}}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{mdframed}
    \begin{theorem}[Reingold]
      There is an $\mathcal{O}(\log n)$-space algorithm producing a universal
      exploration sequence for any regular graph on $n$ vertices.
    \end{theorem}
  \end{mdframed}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[->, bend right] (d1) to (d2);
      \node[rectangle, draw=green, below right=0.3 and 1 of d1]
        {3-Regularity};
      \node[rectangle, draw=blue, below right=1 and 1.2 of d1]
        {Periodic Universal Lemma};
      \node[rectangle, draw=yellow, below right=1.7 and 1 of d1]
        {Closed Walk Lemma};
    \end{tikzpicture}
  \end{center}
  \begin{mdframed}
    \begin{theorem}
      There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
      that moves along a closed walk and either explores the graph or visits at
      least $n$ distinct vertices, for any graph with bounded degree.
    \end{theorem}
  \end{mdframed}
\end{frame}

@@

\begin{frame}
  \frametitle{2. Finding distinct vertices}
  Problem: Indistinguishable vertices $\leadsto$ next vertex \enquote{new}?
  \begin{columns}
    \column{0.5\textwidth}
    Solution:
    \begin{enumerate}
      \item \textcolor<3,8>{magenta}{Execute step of $M_{\text{walk}}$}
      \item \textcolor<4,9>{magenta}{Drop pebble $p_{\text{tmp}}$}
      \item \textcolor<4,9>{magenta}{Store $\Twalk$ in $T'_{\text{walk}}$}
      \item \textcolor<5,10>{magenta}{Restart $M_{\text{walk}}$}
      \item \textcolor<6,11-18>{magenta}{Execute steps of $M_{\text{walk}}$
          until $p_{\text{tmp}}$ is observed}
      \item \textcolor<19-20>{magenta}{If $\Twalk = T'_{\text{walk}}$ the
          vertex is new (increase $\Tid$)}
      \item \textcolor<7>{magenta}{Otherwise start over}
    \end{enumerate}
    \column{0.5\textwidth}
    \input{tikz/disvert}
  \end{columns}
\end{frame}

@@

\begin{frame}
  \frametitle{Lower pebble bound}
  \begin{mdframed}
    \begin{theorem}
      For any constant $\epsilon > 0$, an agent with at most
      $\mathcal{O}((\log n)^{1-\epsilon})$ bits of memory needs at least
      $\Omega(\log\log n)$ distinguishable pebbles for exploring all
      graphs on at most $n$ vertices.
    \end{theorem}
  \end{mdframed}
\end{frame}

@@

\begin{frame}
  \begin{center}
    \Huge Thank you for your attention.
  \end{center}
\end{frame}

>>>>

\usepackage[utf8]{inputenc}
\usepackage{epigraph}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{mdframed}
\usepackage{graphicx}
\usepackage{appendixnumberbeamer}
\usepackage{url}

@@

\begin{frame}
  \frametitle{Agents}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Agent carries set of distinguishable pebbles
      \item Formalisation as pebble machines
        \begin{itemize}
          \item $(Q,F,P,m,\delta,\din,\dout,q_{0})$
          \item $(s,p,m)$-pebble machine
          \item<2-> $\din$: moving onto vertex
          \item<8-> $\dout$: leaving vertex
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \begin{itemize}
      \item<3-> As an agent moves onto a vertex it observes:
        \begin{itemize}
          \item<4-> Port it entered through
          \item<5-> Pebbles it carries
          \item<6-> Pebbles on the vertex
          \item<7-> Degree of vertex
        \end{itemize}
      \item<9-> Computation on a vertex determines:
        \begin{itemize}
          \item<10-> Port to leave
          \item<11-> Pebbles to drop
          \item<12-> Pebbles to carry along
        \end{itemize}
    \end{itemize}
  \end{columns}
\end{frame}

@@

\begin{frame}
  \frametitle{$3$-Regularity}
  \begin{mdframed}[linecolor=purple]
    Establish $3$-Regularity
  \end{mdframed}
  \begin{center}
    \uncover<2->{\resizebox{0.7\textwidth}{!}{\input{tikz/regsim}}}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{mdframed}
    \begin{theorem}[Reingold]
      There is an $\mathcal{O}(\log n)$-space algorithm producing a universal
      exploration sequence for any regular graph on $n$ vertices.
    \end{theorem}
  \end{mdframed}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[->, bend right] (d1) to (d2);
      \node[rectangle, draw=purple, below right=0.3 and 1 of d1]
        {3-Regularity};
      \node[rectangle, draw=blue, below right=1 and 1.2 of d1]
        {Periodic Universal Lemma};
      \node[rectangle, draw=yellow, below right=1.7 and 1 of d1]
        {Closed Walk Lemma};
    \end{tikzpicture}
  \end{center}
  \begin{mdframed}
    \begin{theorem}
      There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
      that moves along a closed walk and either explores the graph or visits at
      least $n$ distinct vertices, for any graph with bounded degree.
    \end{theorem}
  \end{mdframed}
\end{frame}

@@

\begin{frame}
  \frametitle{2. Finding distinct vertices}
  Problem: Indistinguishable vertices $\leadsto$ next vertex \enquote{new}?
  \begin{columns}
    \column{0.5\textwidth}
    Solution:
    \begin{enumerate}
      \item \textcolor<3,8>{magenta}{Execute step of $M_{\text{walk}}$}
      \item \textcolor<4,9>{magenta}{Drop pebble $p_{\text{tmp}}$}
      \item \textcolor<4,9>{magenta}{Store number of steps $M_{\text{walk}}$
        executed}
      \item \textcolor<5,10>{magenta}{Restart $M_{\text{walk}}$}
      \item \textcolor<6,11-18>{magenta}{Execute steps of $M_{\text{walk}}$
          until $p_{\text{tmp}}$ is observed}
      \item \textcolor<19-20>{magenta}{Same amount of steps for
        $M_{\text{walk}}$ to find $p_{\text{tmp}}$ $\rightarrow$ new vertex}
      \item \textcolor<7>{magenta}{Otherwise start over}
    \end{enumerate}
    \column{0.5\textwidth}
    \input{tikz/disvert}
  \end{columns}
\end{frame}

@@

\begin{frame}
  \frametitle{Lower pebble bound}
  \begin{mdframed}
    \begin{theorem}
      For $r\leq p$ and $s\geq 2^{p}$, the number of vertices of an $r$-barrier
      $B_{r}$ for the $s$-state agent $\mathcal{A}$ with $p$ pebbles is bounded
      by $\mathcal{O}(s^{8^{r+1}})$.
    \end{theorem}
  \end{mdframed}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=1.5 of d1] (d2) {};
      \draw[->, bend right] (d1) to (d2);
    \end{tikzpicture}
  \end{center}
  \begin{mdframed}
    \begin{theorem}
      For any constant $\epsilon > 0$, an agent with at most
      $\mathcal{O}((\log n)^{1-\epsilon})$ bits of memory needs at least
      $\Omega(\log\log n)$ distinguishable pebbles for exploring all
      graphs on at most $n$ vertices.
    \end{theorem}
  \end{mdframed}
\end{frame}

@@

\begin{frame}
  \vspace{0.3\textheight}
  \begin{center}
    {\Huge Thank you for your attention.}
  \end{center}
  \vspace{0.35\textheight}
  Slides and text on \url{https://github.com/weltoph/sem-pebbles}.
\end{frame}
